Forecasting hourly hospital admission

### 2021-10-21 paper/paper.Rmd
Update GAMLSS discussion

If we assume that our predictive distribution follows a given parametric distribution, the forecasting task becomes ones of predicting the future values of that distribution's parameters. Generalised Additive Models for Location, Scale and Shape (GAMLSS) are distributional regression models where the parameters of the assumed distribution may be modelled as additive functions of explanatory variables. This provides a powerful and flexible framework for probabilistic forecasting, provided that a suitable distribution and additive structures can be found. In practice, this means experimenting with various distributions and evaluating their suitability using available training data.
Let $y_t$ be the number of attendances in time period $t$ and indicate with $F_t(y_t)$ its predictive cumulative probability distribution. In a distributional regression context, $F_t(y_t)$ is modelled via a parametric model, $F(y_t|\bm \theta_t)$, where $\bm \theta_t$ is an $m$-dimensional vector of parameters. In a GAMLSS framework @Rigby2005 the elements $j=1,...,m$ of $\bm \theta_t$ are modelled via

@@

As our data are counts, the natural starting point is the Poisson distribution, given by
\begin{equation}
  F_t(y_t,\lambda_t) =  \frac{\Gamma(\lfloor y_t + 1  \rfloor,\lambda_t)}{\lfloor y_t \rfloor}
  (\#eq:poissonreg)
\end{equation}
where we initially consider an additive model for $\lambda_t$ of the form
\begin{equation}
  \log(\lambda_t) = \sum_{i=1}^7 \beta_i D_i(t) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad ,
 (\#eq:additivemodel)
\end{equation}
where the functions $H(t)$ and $Y(t)$ return the hour of the day (1--24) and day of the year (1-366) at time $t$, respectively, $D_i(t)$ is a dummy variable for day of the week at $t$, and $C_t$ is the temperature at time $t$. Analysis of this model fit on the training data revealed systematic biases which were identified as relating to public and school holidays. We therefore expand the model to include these effects as follows,
\begin{equation}
  \log(\lambda_t) = \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad ,
 (\#eq:additivemodel_v2)
\end{equation}
where $D^{+}_i(t)$ is a dummy variable for day type including Monday to Sunday, Public Holiday, Christmas Day and New Year's Day, and $S_k(t)$ is a dummy variable for School Holidays.

However, calibration of resulting forecasts on the training data is poor, suggesting that the shape of the Poisson distribution is unsuitable for the present application. In particular, we observed that forecast uncertainty appeared to vary depending on the time of day and possibly other explanatory variables. Therefore, we also consider more flexible two parameter distributions so that we are able to specify additive models for both location and scale parameters, specifically the truncated Normal distribution, with truncation at 0. The resulting density forecasts are given by
\begin{equation}
  F_t(y_t,\mu_t,\sigma_t) =  \frac{\Phi\left( \frac{y_t-\mu_t}{\sigma_t} \right) - \Phi\left( \frac{-y_t}{\sigma_t} \right)}{1 - \Phi\left( \frac{-y_t}{\sigma_t} \right)}
 (\#eq:truncatedn)
\end{equation}
with additive models
\begin{align*}
  \mu_t = &  \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad, \\
  log(\sigma_t) & = \sum_{i=1}^10 D^{+}_i(t) f(H(t))
\end{align*}
for the mean and variance parameters.

Truncated $t$ distribution...

Negative binomial...

>>>>

If we assume that our predictive distribution follows a given parametric distribution, the forecasting task becomes ones of predicting the future values of that distribution's parameters. Generalised Additive Models for Location, Scale and Shape (GAMLSS) are distributional regression models where the parameters of the assumed distribution may be modelled as additive functions of explanatory variables. This provides a powerful and flexible framework for probabilistic forecasting, provided that a suitable distribution and additive structures can be found. In practice, this means experimenting with various distributions and evaluating their suitability using available training data. Here we use this framework to extend the Poisson regression introduced above, and explore alternative distirbutional assumptions.

Let $y_t$ be the number of attendances in time period $t$ and indicate with $F_t(y_t)$ its predictive cumulative probability distribution. In a distributional regression context, $F_t(y_t)$ is modelled via a parametric model, $F(y_t|\bm \theta_t)$, where $\bm \theta_t$ is an $m$-dimensional vector of parameters. In a GAMLSS framework @Rigby2005 the elements $j=1,...,m$ of $\bm \theta_t$ are modelled via

@@

As our data are counts, the natural starting point is the Poisson distribution, given by
\begin{equation}
  F_t(y_t,\lambda_t) =  \frac{\Gamma(\lfloor y_t + 1  \rfloor,\lambda_t)}{\lfloor y_t \rfloor}
  (\#eq:poissonreg)
\end{equation}
<!-- where we initially consider an additive model for $\lambda_t$ of the form -->
<!-- \begin{equation} -->
<!--   \log(\lambda_t) = \sum_{i=1}^7 \beta_i D_i(t) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad , -->
<!--  (\#eq:additivemodel) -->
<!-- \end{equation} -->
<!-- where the functions $H(t)$ and $Y(t)$ return the hour of the day (1--24) and day of the year (1-366) at time $t$, respectively, $D_i(t)$ is a dummy variable for day of the week at $t$, and $C_t$ is the temperature at time $t$. Analysis of this model fit on the training data revealed systematic biases which were identified as relating to public and school holidays. We therefore expand the model to include these effects as follows, -->
<!-- \begin{equation} -->
<!--   \log(\lambda_t) = \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad , -->
<!--  (\#eq:additivemodel_v2) -->
<!-- \end{equation} -->
<!-- where $D^{+}_i(t)$ is a dummy variable for day type including Monday to Sunday, Public Holiday, Christmas Day and New Year's Day, and $S_k(t)$ is a dummy variable for School Holidays. -->
where we consider an additive model for $\lambda_t$ of the form
\begin{equation}
  \log(\lambda_t) = \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad ,
 (\#eq:additivemodel_v2)
\end{equation}
where the functions $H(t)$ and $Y(t)$ return the hour of the day (1--24) and day of the year (1-366) at time $t$, respectively, where $D^{+}_i(t)$ is a dummy variable for day type including Monday to Sunday, Public Holiday, Christmas Day and New Year's Day, $S_k(t)$ is a dummy variable for School Holidays, and $C_t$ is the temperature at time $t$. Public and school holidays were included following experiments on the training data which revealed systematic biases associated with these periods.

However, experiments on the training data reveal that calibration of forecasts based on the Poisson distribution is poor, suggesting that the shape of the distribution is unsuitable for the present application. In particular, we observe that forecast uncertainty appears to vary depending on the time of day and possibly other explanatory variables. Therefore, we consider more flexible, two parameter distributions so that we are able to specify additive models for both location and scale parameters, specifically the truncated Normal distribution, with truncation at 0. The resulting density forecasts are given by
\begin{equation}
  F_t(y_t,\mu_t,\sigma_t) =  \frac{\Phi\left( \frac{y_t-\mu_t}{\sigma_t} \right) - \Phi\left( \frac{-y_t}{\sigma_t} \right)}{1 - \Phi\left( \frac{-y_t}{\sigma_t} \right)}
 (\#eq:truncatedn)
\end{equation}
with additive models
\begin{align*}
  \mu_t = &  \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad, \\
  log(\sigma_t) & = \sum_{i=1}^10 D^{+}_i(t) f(H(t))
\end{align*}
for the mean and variance parameters. We also performed experiments with the truncated $t$ distribution and negative binomial distribution, but these did not results in forecasts as well calibrated as the truncated normal.
